An equivalence of compositional frameworks

### 2018-12-30 EquivCompFrameworks_20180911kc.tex
Continuing work on intro

The second author in a previous work extended...

We could label the edge $e$ something else, for instance, $e^\prime$, and this single edge would another element of $F(N)$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$v_1$};
\node (B) at (1,0) {$v_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$e^\prime$} (B);
\end{tikzpicture}
\]
Denote this second element of $F(N)$ as $d^\prime$. These two graphs constitute distinct isomorphism classes in the symmetric monoidal category $F$Cospan$(\textrm{FinSet})$, as $f=\id_N$ and thus $F(f)=\id_{F(N)}$, but clearly $d \neq d^\prime$. The issue here is that the functor $F$ is decorating each finite set $N$ not only with extra structure, but with extra \emph{stuff}. Moreover, the finite set $N$ is being decorated with \emph{different} stuff in each instance and this difference in stuff is not entirely detected by the function $F(f)$. This phenomenon doesn't occur when the decorations only involve extra structure, and this has been utilized in other works \cite{BFP,BP,Yass}.

One approach to remedying this nuisance is to instead view $F$ as a functor $F \colon \textrm{A} \to \mathbf{Cat}$ and then consider $F(c)$ not as a \emph{set} of decorations on the element $c$ but as a \emph{category} of decorations on the element $c$. Then one can construct a similar symmetric monoidal category similar to Fong's, which will be called $D(H(\mathbb{F}\textnormal{Cospan}(\mathrm{A})))$ for reasons that will be explained later. This symmetric monoidal category will have:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of cospans of $\mathrm{A}$ together with an element of image of the apex under the functor $F$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Except now, given another morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two morphisms are in the same isomorphism class if the following diagrams commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in the category $F(c^\prime)$.
}
\end{enumerate}
Note that the existence of an isomorphism $\iota \colon F(f)(d) \to d^\prime$ is equivalent to the triangle above on the right commuting up to a 2-isomorphism.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (4,0) {$\iota \Swarrow$};
\node (D) at (3,0) {$1$};
\node (E) at (4.5,1) {$F(c)$};
\node (E') at (4.5,-1) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In other words, the isomorphism $\iota$ is a \emph{natural} isomorphism. We will typically just write this as $\iota \colon F(f)(d) \to d^\prime$ to conserve space.

>>>>

In a previous work, the second author extended Fong's theory of decorated cospans to a symmetric monoidal bicategory of decorated cospans \cite{Cour}. Namely, under the same hypotheses used by Fong amounting to a category $\mathsf{A}$ with finite colimits and a symmetric lax monoidal functor $F \colon (\mathsf{A},+,0) \to (\mathsf{Set}, \times, 1)$, there exists a symmetric monoidal bicategory $F\mathbf{Cospan}(\mathsf{A})$ which has:
\begin{enumerate}
\item objects as those of $\mathsf{A}$,
\item morphisms as $F$-decorated cospans in $\mathsf{A}$, which again are pairs
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0) {$F(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(D) edge node [above] {$d$} (E);
\end{tikzpicture}
\]
and
\item 2-morphisms as pairs of commuting diagrams where the left diagram is a map of cospans in $\mathsf{A}$ and the right diagram is a commutative triangle in $\mathsf{Set}$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (B') at (1,-1) {$b^\prime$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(b)$};
\node (E') at (4,-0.5) {$F(b^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
%(B) edge node [right] {$\sim$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
\end{enumerate}
Returning to the previous example where we take the functor $F$ to be the functor $F \colon \mathsf{FinSet} \to \mathsf{Set}$, the situation becomes more dire. Previously, in the ordinary symmetric monoidal category $F$Cospan$(\mathsf{FinSet})$ the following two $F$-decorated cospans resided in distinct isomorphism classes:
\[
\begin{tikzpicture}
  [scale=.8,auto=left]
%\node (D) at (9,11.5) {$1$};
\node (E) at (9,13) {$d \in F(b)$};
\node (E') at (9,10) {$d^\prime \in F(b^\prime)$};
\node (a) at (-2,13) {$1$};
\node (a') at (-2,10) {$1$};
\node (b) at (7,13) {$2$};
\node (b') at (7,10) {$2$};
\node (i) at (-0.5,13.5) {$i$};
\node (o) at (5.5,13.5) {$o$};
\node (i') at (-0.5,10.5) {$i^\prime$};
\node (o') at (5.5,10.5) {$o^\prime$};
%\node (A) at (2.5,11.5) {$\nexists F(f) \Downarrow$};
  \node [style={circle,fill=teal}] (n1) at (1,13) {$v_1$};
  \node[style={circle,fill=teal}] (n2) at (4,13)  {$v_2$};
  \node [style={circle,fill=red}] (n1') at (1,10) {$w_1$};
  \node[style={circle,fill=red}] (n2') at (4,10)  {$w_2$};
\path[->,font=\scriptsize,>=angle 90]
%(D) edge node [above] {$d$} (E)
%(D) edge node [below] {$d^\prime$} (E')
%(E) edge[dashed] node [right] {$\nexists F(f)$} (E')
(a) edge[dashed] node[above] {$$} (n1)
(b) edge[dashed] node[above] {$$} (n2)
(a') edge[dashed] node[below] {$$} (n1')
(b') edge[dashed] node[below] {$$} (n2')
(n1) edge node[above]{$e$} (n2)
(n1') edge node[above]{$e^\prime$} (n2');
\end{tikzpicture}
\]
In the symmetric monoidal bicategory $F\mathbf{Cospan}(\mathsf{FinSet})$, the above two $F$-decorated cospans are two morphisms - with \emph{no 2-morphism between them!} The culprit for this is the same as before - the requirement that the triangle of decorations in $\mathsf{Set}$ commute `on the nose'. In this paper we construct a similar symmetric monoidal bicategory $F$Csp which will remedy this situation, and whose decategorification will remedy the analogous problem for decorated cospan categories above.

It is worth noting that in both examples involving sets and graphs above, we are decorating each finite set with extra \emph{stuff}. This phenomenon does not occur when the decorations only involve extra structure, and this has been utilized in other works \cite{BFP,BP,Yass}.

Our approach is to instead view the functor $F$ as a pseudofunctor $F \colon \textrm{A} \to \mathbf{Cat}$ and then consider $F(a)$ not as a \emph{set} of decorations on the object $a$ but as a \emph{category} of decorations on the element $a$. Then we first construct a symmetric monoidal double category $F \mathbb{C}$sp. This symmetric monoidal double category will have:
\begin{enumerate}
\item{objects as those of $\mathrm{A}$ and}
\item{morphisms as isomorphism classes of cospans of $\mathrm{A}$ together with an element of image of the apex under the functor $F$.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Except now, given another morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$d^\prime \in F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
these two morphisms are in the same isomorphism class if the following diagrams commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,1) {$c$};
\node (B') at (1,-1) {$c^\prime$};
\node (C) at (2,0) {$b$};
\node (D) at (3,0) {$1$};
\node (E) at (4,0.5) {$F(c)$};
\node (E') at (4,-0.5) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i^\prime$} (B')
(C) edge node[below]{$o^\prime$} (B')
(B) edge node [left] {$f$} (B')
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
and there exists an isomorphism $\iota \colon F(f)(d) \to d^\prime$ in the category $F(c^\prime)$.
}
\end{enumerate}
Note that the existence of an isomorphism $\iota \colon F(f)(d) \to d^\prime$ is equivalent to the triangle above on the right commuting up to a 2-isomorphism.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (4,0) {$\iota \Swarrow$};
\node (D) at (3,0) {$1$};
\node (E) at (4.5,1) {$F(c)$};
\node (E') at (4.5,-1) {$F(c^\prime)$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$d$} (E)
(D) edge node [below] {$d^\prime$} (E')
(E) edge node [right] {$F(f)$} (E');
\end{tikzpicture}
\]
In other words, the isomorphism $\iota$ is a \emph{natural} isomorphism. We will typically just write this as $\iota \colon F(f)(d) \to d^\prime$ to conserve space.
